Supplement to ``Local clustering of density upper level sets''

### 2019-01-19 icml/icml_local_density_clustering_supplement.tex
midway through first lemma of pointwise consistency

\newcommand{\diam}{\mathrm{diam}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\defeq}{\overset{\mathrm{def}}{=}}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\1}{\mathbf{1}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\Err}{\mathrm{Err}}
\newcommand{\Log}{\mathrm{Log}}

@@

@@

\begin{lemma}
	\label{lem: consistency_of_conductance_function}
	Fix $0 < t < 1/2$. Under the conditions on $\Csig$ given by Theorem \ref{thm: inverse_mixing_time_lower_bound}, the following statement holds: with probability one, as $n \to \infty$,
	\begin{equation}
	\label{eqn: pointwise_consistency_of_conductance_function}
	\liminf_{n \to \infty} \widetilde{\Phi}_{n,r}(t) \geq \widetilde{\Phi}_{\Pbb,r}(t)
	\end{equation}
	As a consequence, for $M$ and $(t_i)_{i=0}^{M-1}$ defined as in Lemma \ref{lem: continuous_conductance_lower_bound}, we have that
	\begin{equation}
	\label{eqn: consistency_of_stepapprox_conductance_function}
	\liminf_{n \to \infty} \overline{\Phi}_{n,r} \geq \overline{\Phi}_{\Pbb,r}
	\end{equation}
\end{lemma}
We defer the proof of pointwise consistency to Section \ref{subsection: proof_of_pointwise consitency_of_conductance_function}. For now, we show assume  that \eqref{eqn: consistency_of_stepapprox_conductance_function} is immediately implied by \eqref{eqn: pointwise_consistency_of_conductance_function}.

@@

>>>>

\newcommand{\diam}{\mathrm{diam}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\defeq}{\overset{\mathrm{def}}{=}}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\1}{\mathbf{1}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\Err}{\mathrm{Err}}
\newcommand{\Log}{\mathrm{Log}}
\newcommand{\TV}{\mathrm{TV}}

@@

%%% Tilde notation for quantities over the expansion set 
\newcommand{\wn}{\widetilde{n}}
\newcommand{\wX}{\widetilde{\Xbf}}
\newcommand{\wx}{\widetilde{x}}
\newcommand{\wu}{\widetilde{u}}
\newcommand{\wPbb}{\widetilde{\Pbb}}

@@

\begin{lemma}
	\label{lem: consistency_of_conductance_function}
	Fix $0 < t < 1/2$. Under the conditions on $\Csig$ given by Theorem \ref{thm: inverse_mixing_time_lower_bound}, the following statement holds: with probability one, as $n \to \infty$,
	\begin{equation}
	\label{eqn: pointwise_consistency_of_conductance_function}
	\liminf_{n \to \infty} \widetilde{\Phi}_{n,r}(t) \geq \widetilde{\Phi}_{\Pbb,r}(t)
	\end{equation}
	As a consequence, for $M$ and $(t_i)_{i=0}^{M-1}$ defined as in Lemma \ref{lem: continuous_conductance_lower_bound}, we have that
	\begin{equation}
	\label{eqn: consistency_of_stepapprox_conductance_function}
	\liminf_{n \to \infty} \overline{\Phi}_{n,r} \geq \overline{\Phi}_{\Pbb,r}
	\end{equation}
\end{lemma}
We defer the proof of pointwise consistency to Section \ref{subsection: proof_of_pointwise consitency_of_conductance_function}. For now, we show that \eqref{eqn: consistency_of_stepapprox_conductance_function} is immediately implied by \eqref{eqn: pointwise_consistency_of_conductance_function}.

@@

Throughout, let $\wn = \abs{\Csig[\Xbf]}$, $\wX = \Csig[\Xbf] = \set{\wx_1, \ldots, \wx_{\wn}}$, and 
\begin{equation*}
\widetilde{\Pbb}_n := \frac{1}{\wn} \sum_{\wx_i \in \wX} \delta_{\wx_i}
\end{equation*}
be the empirical distribution of $\wX$. For $S \subset \wX$, we will denote $\vol(S; \widetilde{G}_{n,r})$ by $\widetilde{\vol}_{n,r}(S)$, and likewise $\cut(S; \widetilde{G}_{n,r})$ by $\widetilde{\cut}_{n,r}(S)$.

Consider a sequence of sets $(S_{\wn})_{\wn \in \mathbb{N}}$, with $u_{\wn} = \1_{S_{\wn}}$ the characteristic function of $S_{\wn}$. Similarly, for $\Sset \subset \Csig$ let $u = \1_{\Sset}$. 
\begin{lemma}
	If $u_{\wn} \overset{TL^1}{\to} u$,
	\begin{equation*}
	\liminf_{n \to \infty} \frac{\widetilde{\cut}_{n,r}(S_{\wn})}{\wn^2} \geq \widetilde{Q}_{\Pbb,r}(S, \Csig \setminus S) \cdot \left(\int_{\Csig \cap B(x,r)} f(x) \right)
	\end{equation*}
\end{lemma}
\begin{proof}
	We note immediately that $n \to \infty$ implies $\wn \to \infty$ with probability one.

	Now, we can write
	\begin{align*}
	 \frac{\widetilde{\cut}_{n,r}(S_{\wn})}{\wn^2} & = \frac{1}{\wn^2} \sum_{i = 1}^{\wn} \sum_{j = 1}^{\wn} u_{\wn}(\wx_i) \bigl( 1 - u_{\wn}(\wx_j) \1(\norm{\wx_i - \wx_j} \leq r) \bigr)  \\
	& = \int_{\Csig} \left(\int_{\Csig \cap B(x,r)} u_{\wn}(x) \bigl( 1 - u_{\wn}(x') \bigr) d\Pbb_n(x') \right) d\Pbb_n(x) \\
	& = \int_{\Csig} \left(u_{\wn} \circ T_{\wn}(x)\right) \left(\int_{\Csig \cap B(T_{\wn}(x),r)}  1 - u_{\wn} \circ T_{\wn}(x') d\Pbb(x') \right) d\Pbb(x).
	\end{align*}

	Note that
	\begin{align*}
	\lim_{n \to \infty} \nu\bigl(B(T_{\wn}(x),r) \setminus B(x,r)\bigr) & = 0 \\
	\lim_{n \to \infty} \nu\bigl(B(x,r) \setminus B(T_{\wn}(x),r)\bigr) & = 0
	\end{align*}
	and therefore 
	\begin{equation*}
	\int_{\Csig \cap B(T_{\wn}(x),r)}  1 - u_{\wn} \circ T_{\wn}(x') d\Pbb(x') = \int_{\Csig \cap B(x,r)}  1 - u_{\wn} \circ T_{\wn}(x') d\Pbb(x')
	\end{equation*}
	Writing $u_{\wn} \circ T_{\wn}(x') = $
\end{proof}
